---
title: ""
author: ""
date: ""
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: false
    toc_depth: 3
    smooth_scroll: false
    theme: cerulean
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(ggpubr)
library(gridExtra)
library(grid)
library(cowplot)
library(reactable)
library(data.table)
library(reticulate)
library(sciplot)
library(plotly)
library(matrixStats)
library(SummarizedExperiment)
library(RColorBrewer)
library(heatmaply)
library(ggrepel)
library(tidyverse)
library(gprofiler2)
library(patchwork)
library(DESeq2)
library(apeglm)
library(EnsDb.Hsapiens.v79)
library(org.Hs.eg.db)
library(fgsea)
library(msigdbr)
library(compiler)
library(shiny)
library(tippy)
library(msigdbr)
library(GSVA)
library(pals)
library(ggtext)
library(broom)
library(DT)
#load RData
load("~/Documents/Q005841_HNSCC/work/processed_dat/cdat_hnscc.RData")
load("~/Documents/Q005841_HNSCC/work/processed_dat/dds_Dec2021_hnscc.RData")
load("~/Documents/Q005841_HNSCC/work/processed_dat/rld_hnscc_Dec2021.RData")
```

# PCA: only new samples {.tabset .tabset-pills}
## All samples
```{r echo = FALSE, warning=FALSE,message=FALSE,out.width='100%',fig.cap="Top 500 most variably expressed genes used"}
rv <- rowVars(assay(rld))
colnames(rld) <- gsub(pattern = "_2.19.*",replacement = "",colnames(rld)) #removes extra names
select <- order(rv, decreasing=T)[seq_len(min(500,length(rv)))]
pc <- prcomp(t(assay(rld)[select,]))
condition <- cdat_hnscc_final$condition
#cell_line <- cdat_hnscc_final$cell_line
scores <- data.frame(pc$x, condition)
percentage <- round(pc$sdev / sum(pc$sdev) * 100, 2)
percentage <- paste( colnames(scores), "(", paste( as.character(percentage), "%", ")", sep="") )
#plot
plot_ly(scores,x=scores$PC1,y=scores$PC2,text=rownames(scores),mode="markers",color=factor(condition),marker=list(size=11),colors = "Paired",type="scatter") %>% layout(title="",   xaxis = list(title = percentage[1]),yaxis = list(title = percentage[2]))
```

## Cal27
```{r echo = FALSE, warning=FALSE,message=FALSE,out.width='100%'}
rld.cal27 <- rld[ , rld$cell_line %in% c("Cal27")]
cal27_cdat <- cdat_hnscc_final %>%
  dplyr::filter(cell_line == "Cal27")
rv <- rowVars(assay(rld.cal27))
select <- order(rv, decreasing=T)[seq_len(min(500,length(rv)))]
pc <- prcomp(t(assay(rld.cal27)[select,]))
condition <- cal27_cdat$condition
#cell_line <- cdat_hnscc_final$cell_line
scores <- data.frame(pc$x, condition)
percentage <- round(pc$sdev / sum(pc$sdev) * 100, 2)
percentage <- paste( colnames(scores), "(", paste( as.character(percentage), "%", ")", sep="") )
#plot
plot_ly(scores,x=scores$PC1,y=scores$PC2,text=rownames(scores),mode="markers",color=factor(condition),marker=list(size=11),colors = "Paired",type="scatter") %>% layout(title="",   xaxis = list(title = percentage[1]),yaxis = list(title = percentage[2]))
```

## Detroit
```{r echo = FALSE, warning=FALSE,message=FALSE,out.width='100%'}
rld.detroit <- rld[ , rld$cell_line %in% c("Detroit")]
detroit_cdat <- cdat_hnscc_final %>%
  dplyr::filter(cell_line == "Detroit")
rv <- rowVars(assay(rld.detroit))
select <- order(rv, decreasing=T)[seq_len(min(500,length(rv)))]
pc <- prcomp(t(assay(rld.detroit[select,])))
condition <- detroit_cdat$condition
#cell_line <- cdat_hnscc_final$cell_line
scores <- data.frame(pc$x, condition)
percentage <- round(pc$sdev / sum(pc$sdev) * 100, 2)
percentage <- paste( colnames(scores), "(", paste( as.character(percentage), "%", ")", sep="") )
#plot
plot_ly(scores,x=scores$PC1,y=scores$PC2,text=rownames(scores),mode="markers",color=factor(condition),marker=list(size=11),colors = "Paired",type="scatter") %>% layout(title="",   xaxis = list(title = percentage[1]),yaxis = list(title = percentage[2]))
```
## FaDu & SKN3
```{r echo = FALSE, warning=FALSE,message=FALSE,out.width='100%'}
rld.fadu.skn3 <- rld[ , rld$cell_line %in% c("FaDu","SKN3")]
fadu_skn3_cdat <- cdat_hnscc_final %>%
  dplyr::filter(cell_line == "SKN3" | cell_line == "FaDu")
rv <- rowVars(assay(rld.fadu.skn3))
select <- order(rv, decreasing=T)[seq_len(min(500,length(rv)))]
pc <- prcomp(t(assay(rld.fadu.skn3[select,])))
condition <- fadu_skn3_cdat$condition
#cell_line <- cdat_hnscc_final$cell_line
scores <- data.frame(pc$x, condition)
percentage <- round(pc$sdev / sum(pc$sdev) * 100, 2)
percentage <- paste( colnames(scores), "(", paste( as.character(percentage), "%", ")", sep="") )
#plot
plot_ly(scores,x=scores$PC1,y=scores$PC2,text=rownames(scores),mode="markers",color=factor(condition),marker=list(size=11),colors = "Paired",type="scatter") %>% layout(title="",   xaxis = list(title = percentage[1]),yaxis = list(title = percentage[2]))
```
## {-}

# Correlation Heatmaps: only new samples {.tabset .tabset-pills}
## All samples
```{r echo = FALSE, warning=FALSE,message=FALSE,out.width='100%'}
rv <- rowVars(assay(rld))
select <- order(rv, decreasing=T)[seq_len(min(500,length(rv)))]
#sampleDists <- stats::dist(t(assay(rld[select])),method = "euclidean")
#sampleDistMatrix <- as.matrix(sampleDists)
#colors <- colorRampPalette( rev(brewer.pal(9, "YlOrRd")) )(255)
#heatmaply(sampleDistMatrix,
        #  clustering_distance_rows=sampleDists,
       #   clustering_distance_cols=sampleDists,
       #   col=colors,showticklabels = FALSE,key.title = "Euclidean distance",row_dend_left = TRUE,dend_hoverinfo #= FALSE)

heatmaply_cor(
  cor(assay(rld[select,])),showticklabels = FALSE,key.title = "",row_dend_left = TRUE)
```
## Cal27
```{r echo = FALSE, warning=FALSE,message=FALSE,out.width='100%'}
rv <- rowVars(assay(rld.cal27))
select <- order(rv, decreasing=T)[seq_len(min(500,length(rv)))]
heatmaply_cor(
  cor(assay(rld.cal27[select,])),showticklabels = FALSE,key.title = "",row_dend_left = TRUE,dend_hoverinfo = FALSE
)
```

## Detroit
```{r echo = FALSE, warning=FALSE,message=FALSE,out.width='100%'}
rv <- rowVars(assay(rld.detroit))
select <- order(rv, decreasing=T)[seq_len(min(500,length(rv)))]
heatmaply_cor(
  cor(assay(rld.detroit[select,])),showticklabels = FALSE,key.title = "",row_dend_left = TRUE,dend_hoverinfo = FALSE
)
```

## FaDu & SKN3
```{r echo = FALSE, warning=FALSE,message=FALSE,out.width='100%'}
rv <- rowVars(assay(rld.fadu.skn3))
select <- order(rv, decreasing=T)[seq_len(min(500,length(rv)))]
heatmaply_cor(
  cor(assay(rld.fadu.skn3[select,])),showticklabels = FALSE,key.title = "",row_dend_left = TRUE,dend_hoverinfo = FALSE
)
```

## {-}

# Hierarchical clustering: only new samples {.tabset .tabset-pills}
## All samples
```{r echo = FALSE, warning=FALSE,message=FALSE,out.width='100%'}
rv <- rowVars(assay(rld))
select <- order(rv, decreasing=T)[seq_len(min(500,length(rv)))]
gplots::heatmap.2(assay(rld)[select,], margins = c(10, 6),trace = "none", col = viridis(100),srtCol = 70,cexCol = 0.8)
#heatmaply(assay(rld)[select,],seriate = "mean",row_dend_left=TRUE,plot_method="plotly",hide_colorbar = TRUE)
```

## Cal27
```{r echo = FALSE, warning=FALSE,message=FALSE,out.width='100%'}
rv <- rowVars(assay(rld.cal27))
select <- order(rv, decreasing=T)[seq_len(min(500,length(rv)))]
gplots::heatmap.2(assay(rld.cal27)[select,],  margins = c(10, 6),trace = "none", col = viridis(100),srtCol = 70,cexCol = 0.8)
#heatmaply(assay(rld)[select,],seriate = "mean",row_dend_left=TRUE,plot_method="plotly",hide_colorbar = TRUE)
```

## Detroit
```{r echo = FALSE, warning=FALSE,message=FALSE,out.width='100%'}
rv <- rowVars(assay(rld.detroit))
select <- order(rv, decreasing=T)[seq_len(min(500,length(rv)))]
gplots::heatmap.2(assay(rld.detroit)[select,],margins = c(10, 6),trace = "none", col = viridis(100),srtCol = 70,cexCol = 0.8)
#heatmaply(assay(rld)[select,],seriate = "mean",row_dend_left=TRUE,plot_method="plotly",hide_colorbar = TRUE)
```

## FaDu & SKN3
```{r echo = FALSE, warning=FALSE,message=FALSE,out.width='100%'}
rv <- rowVars(assay(rld.fadu.skn3))
select <- order(rv, decreasing=T)[seq_len(min(500,length(rv)))]
gplots::heatmap.2(assay(rld.fadu.skn3)[select,], margins = c(10, 6),trace = "none", col = viridis(100),srtCol = 70,cexCol = 0.8)
#heatmaply(assay(rld)[select,],seriate = "mean",row_dend_left=TRUE,plot_method="plotly",hide_colorbar = TRUE)
```
# {-}

# PCA: old and new HNSCC samples {.tabset .tabset-pills}
## All samples
```{r echo = FALSE, warning=FALSE,message=FALSE,out.width='100%',fig.cap="Top 500 most variably expressed genes used; Circles = new batch; Squares = old batch"}
load("~/Documents/Q005841_HNSCC/data/oldNewHNSCC_dds.RData")
load("~/Documents/Q005841_HNSCC/data/oldnewHNSCC_vst.RData")
load("~/Documents/Q005841_HNSCC/data/oldNew_mdat.RData")
rv <- rowVars(assay(vst))
select <- order(rv, decreasing=T)[seq_len(min(500,length(rv)))]
pc <- prcomp(t(assay(vst)[select,]))
condition <- hnscc_oldNew_mdat$condition
batch <- hnscc_oldNew_mdat$batch
#cell_line <- hnscc_oldNew_mdat$cell_line
scores <- data.frame(pc$x, condition, batch)
percentage <- round(pc$sdev / sum(pc$sdev) * 100, 2)
percentage <- paste( colnames(scores), "(", paste( as.character(percentage), "%", ")", sep="") )
#plot
# text=rownames(scores),mode="markers",color=factor(condition),marker=list(size=11),colors = "Paired",type="scatter",symbol=factor(batch)
symbols <- c('circle','square')
colors <- c('red','blue','green','black','purple','orange','gray','brown','yellowgreen','sienna1','olivedrab','maroon2','lightsteelblue','lightsalmon2','lightpink1','hotpink','chartreuse','azure','tan2','violet','turquoise1','aquamarine')
plot_ly(scores,x=scores$PC1,y=scores$PC2,z=scores$PC3,color=factor(condition), symbol = factor(batch), symbols = symbols, colors = colors, hoverinfo = 'text', text = ~paste('</br>',rownames(scores),'</br>', condition,'</br>',batch)) %>% layout(scene = list(xaxis = list(title = percentage[1]),yaxis = list(title = percentage[2]), zaxis = list(title = percentage[3])))
```

## Cal27
```{r echo = FALSE, warning=FALSE,message=FALSE,out.width='100%'}
vst.cal27 <- vst[ , vst$cell_line %in% c("Cal27")]
cal27_mdat <- hnscc_oldNew_mdat %>%
  dplyr::filter(cell_line == "Cal27")
rv <- rowVars(assay(vst.cal27))
select <- order(rv, decreasing=T)[seq_len(min(500,length(rv)))]
pc <- prcomp(t(assay(vst.cal27)[select,]))
condition <- cal27_mdat$condition
batch <- cal27_mdat$batch
#cell_line <- hnscc_oldNew_mdat$cell_line
scores <- data.frame(pc$x, condition, batch)
percentage <- round(pc$sdev / sum(pc$sdev) * 100, 2)
percentage <- paste( colnames(scores), "(", paste( as.character(percentage), "%", ")", sep="") )
#2d plot
#plot_ly(scores,x=scores$PC1,y=scores$PC2,text=rownames(scores),mode="markers",color=factor(condition),marker=list(size=11),colors = "Paired",type="scatter",symbol=factor(batch)) %>% layout(title="",   xaxis = list(title = percentage[1]),yaxis = list(title = percentage[2]))
#3d plot
symbols <- c('circle','square')
colors <- c('red','blue','green','black','orange','gray','sienna4','yellowgreen','seagreen','olivedrab','maroon2','lightsteelblue','lightsalmon2','cadetblue','hotpink','chartreuse','azure','darkgoldenrod1','turquoise1','aquamarine','midnightblue','slateblue')
plot_ly(scores,x=scores$PC1,y=scores$PC2,z=scores$PC3,color=factor(condition), symbol = factor(batch), symbols = symbols, colors = colors, hoverinfo = 'text', text = ~paste('</br>',rownames(scores),'</br>', condition,'</br>',batch)) %>% layout(scene = list(xaxis = list(title = percentage[1]),yaxis = list(title = percentage[2]), zaxis = list(title = percentage[3])))
```

## Detroit
```{r echo = FALSE, warning=FALSE,message=FALSE,out.width='100%'}
vst.detroit <- vst[ , vst$cell_line %in% c("Detroit")]
detroit_cdat <- hnscc_oldNew_mdat %>%
  dplyr::filter(cell_line == "Detroit")
rv <- rowVars(assay(vst.detroit))
select <- order(rv, decreasing=T)[seq_len(min(500,length(rv)))]
pc <- prcomp(t(assay(vst.detroit[select,])))
condition <- detroit_cdat$condition
batch <- detroit_cdat$batch
#cell_line <- hnscc_oldNew_mdat$cell_line
scores <- data.frame(pc$x, condition,batch)
percentage <- round(pc$sdev / sum(pc$sdev) * 100, 2)
percentage <- paste( colnames(scores), "(", paste( as.character(percentage), "%", ")", sep="") )
#2d plot
#plot_ly(scores,x=scores$PC1,y=scores$PC2,text=rownames(scores),mode="markers",color=factor(condition),marker=list(size=11),colors = "Paired",type="scatter",symbol=factor(batch)) %>% layout(title="",   xaxis = list(title = percentage[1]),yaxis = list(title = percentage[2]))
#3d plot
symbols <- c('circle','square')
colors <- c('red','blue','green','black','orange','gray','sienna4','yellowgreen','seagreen','olivedrab','maroon2','lightsteelblue','lightsalmon2','cadetblue','hotpink','chartreuse','azure','darkgoldenrod1','turquoise1','aquamarine','midnightblue','slateblue')
plot_ly(scores,x=scores$PC1,y=scores$PC2,z=scores$PC3,color=factor(condition), symbol = factor(batch), symbols = symbols, colors = colors, hoverinfo = 'text', text = ~paste('</br>',rownames(scores),'</br>', condition,'</br>',batch)) %>% layout(scene = list(xaxis = list(title = percentage[1]),yaxis = list(title = percentage[2]), zaxis = list(title = percentage[3])))
```
## FaDu
```{r echo = FALSE, warning=FALSE,message=FALSE,out.width='100%'}
vst.fadu.skn3 <- vst[ , vst$cell_line %in% c("FaDu")]
fadu_skn3_cdat <- hnscc_oldNew_mdat %>%
  dplyr::filter(cell_line == "FaDu")
rv <- rowVars(assay(vst.fadu.skn3))
select <- order(rv, decreasing=T)[seq_len(min(500,length(rv)))]
pc <- prcomp(t(assay(vst.fadu.skn3[select,])))
condition <- fadu_skn3_cdat$condition
batch <- fadu_skn3_cdat$batch
#cell_line <- hnscc_oldNew_mdat$cell_line
scores <- data.frame(pc$x, condition,batch)
percentage <- round(pc$sdev / sum(pc$sdev) * 100, 2)
percentage <- paste( colnames(scores), "(", paste( as.character(percentage), "%", ")", sep="") )
#2d plot
#plot_ly(scores,x=scores$PC1,y=scores$PC2,text=rownames(scores),mode="markers",color=factor(condition),marker=list(size=11),colors = "Paired",type="scatter", symbol=factor(batch)) %>% layout(title="",   xaxis = list(title = percentage[1]),yaxis = list(title = percentage[2]))
#3d plot
symbols <- c('circle','square')
colors <- c('red','blue','green','black','orange','gray','sienna4','yellowgreen','seagreen','olivedrab','maroon2','lightsteelblue','lightsalmon2','cadetblue','hotpink','chartreuse','azure','darkgoldenrod1','turquoise1','aquamarine','midnightblue','slateblue')
plot_ly(scores,x=scores$PC1,y=scores$PC2,z=scores$PC3,color=factor(condition), symbol = factor(batch), symbols = symbols, colors = colors, hoverinfo = 'text', text = ~paste('</br>',rownames(scores),'</br>', condition,'</br>',batch)) %>% layout(scene = list(xaxis = list(title = percentage[1]),yaxis = list(title = percentage[2]), zaxis = list(title = percentage[3])))
```
## {-}
# Hierarchical clustering: old and new HNSCC samples {.tabset .tabset-pills}
## All samples
```{r echo = FALSE, warning=FALSE,message=FALSE,out.width='100%',fig.cap="Blue = old batch; Red = new batch"}
rv <- rowVars(assay(vst))
select <- order(rv, decreasing=T)[seq_len(min(500,length(rv)))]
#
cc_all <- 1:52
cc_all[1:27] <- "red"
cc_all[28:52] <- "blue"
#
gplots::heatmap.2(assay(vst)[select,], margins = c(10, 6),trace = "none", col = viridis(100),srtCol = 70,cexCol = 0.5,colCol = cc_all)
#heatmaply(assay(vst)[select,],seriate = "mean",row_dend_left=TRUE,plot_method="plotly",hide_colorbar = TRUE)
```
## Cal27
```{r echo = FALSE, warning=FALSE,message=FALSE,out.width='100%',fig.cap="Blue = old batch; Red = new batch"}
rv <- rowVars(assay(vst.cal27))
select <- order(rv, decreasing=T)[seq_len(min(500,length(rv)))]
#
cc_cal27 <- 1:24
cc_cal27[1:15] <- "red"
cc_cal27[16:24] <- "blue"
#
gplots::heatmap.2(assay(vst.cal27)[select,],  margins = c(10, 6),trace = "none", col = viridis(100),srtCol = 70,cexCol = 0.8,colCol = cc_cal27)
#heatmaply(assay(vst)[select,],seriate = "mean",row_dend_left=TRUE,plot_method="plotly",hide_colorbar = TRUE)
```
## Detroit
```{r echo = FALSE, warning=FALSE,message=FALSE,out.width='100%',fig.cap="Blue = old batch; Red = new batch"}
rv <- rowVars(assay(vst.detroit))
select <- order(rv, decreasing=T)[seq_len(min(500,length(rv)))]
#
cc_Det <- 1:17
cc_Det[1:6] <- "red"
cc_Det[7:17] <- "blue"
#
gplots::heatmap.2(assay(vst.detroit)[select,],margins = c(10, 6),trace = "none", col = viridis(100),srtCol = 70,cexCol = 0.8,colCol = cc_Det)
#heatmaply(assay(vst)[select,],seriate = "mean",row_dend_left=TRUE,plot_method="plotly",hide_colorbar = TRUE)
```
## FaDu
```{r echo = FALSE, warning=FALSE,message=FALSE,out.width='100%',fig.cap="Blue = old batch; Red = new batch"}
rv <- rowVars(assay(vst.fadu.skn3))
select <- order(rv, decreasing=T)[seq_len(min(500,length(rv)))]
#
cc_fad <- 1:6
cc_fad[1:4] <- "red"
cc_fad[5:6] <- "blue"

check <- assay(vst.fadu.skn3)[select,]
#
gplots::heatmap.2(assay(vst.fadu.skn3)[select,], margins = c(10, 6),trace = "none", col = viridis(100),srtCol = 70,cexCol = 0.8,colCol = cc_fad)
#heatmaply(assay(vst)[select,],seriate = "mean",row_dend_left=TRUE,plot_method="plotly",hide_colorbar = TRUE)
```